---
title: "linear_models"
author: "Xiaoyu Wu"
date: "2023-11-05"
output: github_document
---

```{r}
library(tidyverse)
library(p8105.datasets)

set.seed(1)
```

## Model fitting
#### Import data
```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>% 
  select(price, stars, borough, neighborhood, room_type)
```
#### fit a model 
```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
# stars is a numeric predictor and borough is a categorical predictor. 
# different boroughs have paralleled lines 
```

```{r}
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```
Let's look at the result 
```{r}
summary(fit)
summary(fit)$coef
coef(fit)
residuals(fit)
fitted.values(fit)
```
## Tidying output
Let's look at the result better 
```{r}
fit %>% 
  broom::glance()
```

```{r}
fit %>% 
  broom::tidy()
```

```{r}
fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(term = str_replace(term, "^borough", "Borough: ")) %>% 
  knitr::kable(digits = 3)
```
## Be in control of factors 
```{r}
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough),
# in order of frequency of borough 
    room_type = fct_infreq(room_type))

fit = lm(price ~ stars + borough, data = nyc_airbnb)
```
Look at that plot again
```{r}
nyc_airbnb |> 
  ggplot(aes(x=stars,y=price,color=borough))+
  geom_point()
```

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
broom::tidy(fit)
broom::glance(fit)
```


## Diagnostics
```{r}
modelr::add_residuals(nyc_airbnb, fit)
# ask for data and model fit in the brackets  
```

```{r}
modelr::add_predictions(nyc_airbnb, fit)
```

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = borough, y = resid)) + geom_violin()+
  ylim(-500,1500)
```

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = stars, y = resid)) + 
  geom_point() +
  facet_wrap(.~borough)
  
```

## Hypothesis testing
This does t-test by default
```{r}
fit |> 
  broom::tidy()
```

what about significance of `borough` 
```{r}
fit_null = lm(price ~ stars, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough, data = nyc_airbnb)
```

```{r}
anova(fit_null, fit_alt) %>% 
  broom::tidy()
```

## Nesting data, fit models 
This is pretty formal and also complex 
```{r}
nyc_airbnb %>% 
  lm(price ~ stars * borough + room_type * borough, data = .) %>% 
  broom::tidy() %>% 
  knitr::kable(digits = 3)
```
This is more exploratory and easier to understand 
```{r}
nest_lm_res =
  nyc_airbnb %>% 
  nest(data = -borough) %>% 
# nest everything except borough 
  mutate(
    models = map(.x=data, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
# add a result column 
  select(-data, -models) %>% 
  unnest(results) %>%
  filter(term!="(Intercept)")
```

```{r}
nest_lm_res %>% 
  select(borough, term, estimate) %>% 
  mutate(term = fct_inorder(term)) %>% 
  pivot_wider(
    names_from = term, values_from = estimate) %>% 
  knitr::kable(digits = 3)
```
Let's nest even more 
```{r}
manhattan_airbnb =
  nyc_airbnb %>% 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb %>% 
  nest(data = -neighborhood) %>% 
  mutate(
    models = map(data, ~lm(price ~ stars + room_type, data = .x)),
    results = map(models, broom::tidy)) %>% 
  select(-data, -models) %>% 
  unnest(results) 
```

```{r}
manhattan_nest_lm_res %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

## Binary outcomes
```{r}
baltimore_df = 
  read_csv("./homicide-data.csv") %>% 
  filter(city == "Baltimore") %>% 
  mutate(
    resolved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age),
    victim_race = fct_relevel(victim_race, "White")) %>% 
  select(resolved, victim_age, victim_race, victim_sex)
```

```{r}
fit_logistic = 
  baltimore_df %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) 
```

```{r}
fit_logistic %>% 
  broom::tidy() %>% 
  mutate(OR = exp(estimate)) %>%
  select(term, log_OR = estimate, OR, p.value) %>% 
  knitr::kable(digits = 3)
```

```{r}
baltimore_df %>% 
  modelr::add_predictions(fit_logistic) %>% 
  mutate(fitted_prob = boot::inv.logit(pred))
```

